The imprint of global magma oceans on exoplanet demographics

### 2023-01-20 src/tex/ms.tex
discuss CHEOPS and Ariel

\subsection{Detecting the runaway greenhouse transition with current and planned exoplanet missions}
\todo{Introduce survey simulation(s). , Cheops, Ariel, LIFE (reach out to team; Hendrik Linz produced a LIFE Technology Overview document)?, Nautilus\\ What are threshold missions that are able to detect this signal? =>X meter working for Y years, or X' meter working for Y' years => conclusion could be that there is interesting science to do with intermediate size telescope.}

@@

\subsubsection{PLATO}
\plato\ (PLAnetary Transits and Oscillation of stars) is an ESA mission designed to characterize terrestrial planets in the habitable zones of Sun-like stars, a goal to be achieved via long-term high-precision photometric monitoring of a sample of bright stars~\citep{Rauer2016}.
The \plato\ team has released an estimate on the number of exoplanets that will be characterized in the course of the main survey mission (Matuszewski, priv. comm.).
In the FGK sample of the mission, a minimum of 500 Earth-sized (\SIrange{0.8}{1.25}{\rEarth}) planets are expected to be detected.
\plato's definition study report~\citep{plato2017} further states precision requirements for planet radii~(\SI{3}{\percent}), planet masses through radial velocity follow-up~(\SI{10}{\percent}), and stellar masses, radii, and ages~(\SI{10}{\percent}).
We base our assessment of \plato's diagnostic power on these estimates.

@@

\subsection{Statistical power of different mission designs}\label{sec:statpower_missions}
\subsubsection{Additional planet mass measurements}
The situation differs when additional information beyond planet radii is available for the characterized planet population.
Since the planet bulk density $\rho \propto R^{-3}$, the signal is stronger when measured in bulk density instead of radius:
With an optimistic choice of geophysical parameters (see above), the average measured radius change is \SI{+15}{\percent} whereas the average density change is \SI{-33}{\percent}.
Constraining bulk densities requires planetary mass measurements.
We now examine the value of a follow-up campaign to constrain planet masses with respect to the detectability of the runaway greenhouse threshold.

@@

The key findings of our work are as follows:
\begin{enumerate}
    \item The predicted tipping point in planetary climates causes a discontinuity in the radius and density distribution of small (DEFINITION) exoplanets.
    \item The demographic imprint of the runaway greenhouse threshold should be detectable with transit measurements of a large enough sample and sufficient precision.
    \item We find that the planned \plato\ transit survey will provide a sufficient sample to test the runaway greenhouse hypothesis.
    \item The diagnostic power of transit missions in testing this hypothesis can be maximized by ...
    \item A detection of the feature will constrain the water inventory of rocky exoplanets, providing important context in the assessment of their habitability.
\end{enumerate}

>>>>

\subsection{Detecting the runaway greenhouse transition with current and planned exoplanet missions}
\todo{Introduce survey simulation(s). PLATO, (Ariel, LIFE(reach out to team; Hendrik Linz produced a LIFE Technology Overview document)) ?, Nautilus\\ What are threshold missions that are able to detect this signal? =>X meter working for Y years, or X' meter working for Y' years => conclusion could be that there is interesting science to do with intermediate size telescope.}

@@

%\subsubsection{PLATO}
\plato\ (PLAnetary Transits and Oscillation of stars) is an ESA mission designed to characterize terrestrial planets in the habitable zones of Sun-like stars, a goal to be achieved via long-term high-precision photometric monitoring of a sample of bright stars~\citep{Rauer2016}.
The \plato\ team has released an estimate on the number of exoplanets that will be characterized in the course of the main survey mission (Matuszewski, priv. comm.).
In the FGK sample of the mission, a minimum of 500 Earth-sized (\SIrange{0.8}{1.25}{\rEarth}) planets are expected to be detected.
\plato's definition study report~\citep{plato2017} further states precision requirements for planet radii~(\SI{3}{\percent}), planet masses through radial velocity follow-up~(\SI{10}{\percent}), and stellar masses, radii, and ages~(\SI{10}{\percent}).
We base our assessment of \plato's diagnostic power on these estimates.

@@

What other planned missions are suited to test the runaway greenhouse hypothesis?
The ongoing \cheops\ mission was designed as a follow-up mission to search for transits of planets discovered with other techniques, in particular with radial velocity measurements~\citep{Benz2021}.
As such, it will provide precise radius constraints on a sample of small planets; however, only a small number of planets with periods \SI{> 50}{\day} are being observed.
This largely limits \cheops' coverage to planets within the runaway greenhouse regime, preventing a detection of the transition.

As \cheops, \ariel~\citep{Puig2016} will be a follow-up mission that is not designed to provide a large number of new radius measurements.
\ariel's primary targets are larger planets in the range of sub-Neptune to Jupiter-like planets.
We thus do not expect a significant contribution to exploring the inner edge of the habitable zone for Earth-sized planets.

\subsection{Statistical power of different mission designs}\label{sec:statpower_missions}
\subsubsection{Additional planet mass measurements}
The dependencies of a detection change when additional information beyond planet radii is available for the characterized planet population.
Since the planet bulk density $\rho \propto R^{-3}$, the signal is stronger when measured in bulk density instead of radius:
With an optimistic choice of geophysical parameters (see above), the average measured radius change is \SI{+15}{\percent} whereas the average density change is \SI{-33}{\percent}.
Constraining bulk densities requires planetary mass measurements.
We now examine the value of a follow-up campaign to constrain planet masses with respect to the detectability of the runaway greenhouse threshold.

@@

The key findings of our work are as follows:
\begin{enumerate}
    \item The predicted tipping point in planetary climates causes a discontinuity in the radius and density distribution of small exoplanets.
    \item The demographic imprint of the runaway greenhouse threshold should be detectable with transit measurements of a large enough sample and sufficient precision.
    \item We find that the planned \plato\ transit survey will provide a sufficient sample to test the runaway greenhouse hypothesis.
    \item The diagnostic power of transit missions in testing this hypothesis can be maximized by ...
    \item A detection of the feature will constrain the water inventory of rocky exoplanets, providing important context in the assessment of their habitability.
\end{enumerate}
